Equalizer\\Programming and User Guide

### 2012-06-25 Developer/ProgrammingGuide/paper.tex
Programming Guide update for 1.4, part I

\date{
  \vspace{14cm}
  \begin{textblock}{}(0,4.5)
    \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{textblock}
      {\Large The official reference for developing and
        deploying parallel, scalable OpenGL\texttrademark\ applications using
        the Equalizer parallel rendering framework}\\\vspace{1cm}
      Version 1.11.1 for Equalizer 1.2+git\\[\medskipamount]
  \today
}

\newcommand{\tm}{\texttrademark~}
\newcommand{\rc}{\raise 1ex\hbox{{\tiny\textregistered}}~}
\newcommand{\fig}[1]{Figure~\ref{#1}}
\newcommand{\sref}[1]{Section~\ref{#1}}
\newcommand{\aref}[1]{Appendix~\ref{#1}}
\newcommand{\link}[1]{\htmladdnormallink{#1}{#1}}
\indentcaption{2em}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 17, 2012 & Revision for Equalizer 1.2\\
    1.11.1  & Apr 10, 2012 & Minor revision\\
    \hline
  \end{tabularx}}
\clearpage

@@

@@

A view is derived from \textsf{eq::Frustum} (\fig{fViewUML}), and the
application process can set the wall or projection parameters at
runtime. For a description of wall and projection parameters please refer
to \sref{sFrustum}. The new data will be effective for the next frame.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
The default Equalizer event handling is using the view API to maintain
the aspect ratio of destination channels after a window resize. Without
updating the wall or projection description, the rendering would become
distorted.

@@

>>>>

\date{
  \vspace{14cm}
  \begin{textblock}{}(0,4.5)
    \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{textblock}
      {\Large The official reference for developing and
        deploying parallel, scalable OpenGL\texttrademark\ applications using
        the Equalizer parallel rendering framework}\\\vspace{1cm}
      Version 1.11.2 for Equalizer 1.4-beta\\[\medskipamount]
  \today
}

\newcommand{\tm}{\texttrademark~}
\newcommand{\rc}{\raise 1ex\hbox{{\tiny\textregistered}}~}
\newcommand{\fig}[1]{Figure~\ref{#1}}
\newcommand{\sref}[1]{Section~\ref{#1}}
\newcommand{\aref}[1]{Appendix~\ref{#1}}
\newcommand{\link}[1]{\htmladdnormallink{#1}{#1}}
\indentcaption{2em}
\setcounter{tocdepth}{2}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 17, 2012 & Revision for Equalizer 1.2\\
    1.11.1  & Apr 10, 2012 & Minor revision\\
    1.11.2  & Jun 25, 2012 & Added \sref{sScaling} and \sref{sROI}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\subsubsection{\label{sScaling}Application-specific Scaling}

All Equalizer units in the configuration file and API are configured in
meters. This convention allows the reuse of configurations across multiple
applications. When visualizing real-world objects, e.g., for architectural
visualizations, it guarantees that they appear realistic and full immersion
into the visualization can be achieved.

Certain applications want to visualize objects in immersive environments in a
scale different from their natural size, e.g., astronomical simulations. Each
model, and therefore each view, might have a different scale. Applications can
declare this scale as part of the \textsf{eq::View}, which will be applied to
the virtual environment by Equalizer. Common metric model scales are provided as
constants.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
A view is derived from \textsf{eq::Frustum} (\fig{fViewUML}), and the
application process can set the wall or projection parameters at
runtime. For a description of wall and projection parameters please refer
to \sref{sFrustum}. The new data will be effective for the next frame.

@@

The default Equalizer event handling is using the view API to maintain
the aspect ratio of destination channels after a window resize. Without
updating the wall or projection description, the rendering would become
distorted.

@@

\subsection{\label{sROI}Region of Interest}

\subsubsection{Motivation}
Regions of interest (ROI) define which part of the framebuffer was updated by
the application. They allow Equalizer to perform important optimizations during
scalable rendering.

\begin{wrapfigure}{r}{.618\textwidth}
  \subfigure[]{\label{fROIDB2Way}
    \includegraphics[width=.29\textwidth]{images/roiDB2way.png}
  }\hfil
  \subfigure[]{\label{fROIDB4Way}
    \includegraphics[width=.29\textwidth]{images/roiDB4way.png}
  }%
  {\caption{\label{fROIDB}ROI for a two-way (a) and four-way DB (b) compound}}
\end{wrapfigure}
The first optimization concerns the compositing operation. The declared ROI
allows Equalizer to restrict the readback, compression, network transmission and
assembly to only the relevant pixels.

This is particularly important for DB compounds, which typically cover less
screen space on each channel as resources are added to the
decomposition. \fig{fROIDB} illustrates this for a two-way and four-way
decomposition. Declaring the ROI alleviates the increasing compositing cost as
resources are added to a DB compound.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/roi}
  {\caption{\label{fROI}ROI for 2D load balancing}}
\end{wrapfigure}
For 2D compounds, the same optimization is applied, but has a smaller impact on
the compositing performance. \fig{fROI} illustrates the ROI for a four-way 2D
compound.

The second optimization is important for load-balanced 2D compounds. The 2D
load equalizer uses internal timing statistics to form a 2D load grid for
computing the split of the next frame. Without ROI, an even load within each
channel has to be assumed. With ROI, an even load within the declared, smaller
region of interest can be used. This causes a smaller error in the load
estimation, and therefore a more accurate load prediction.

\fig{fROI} illustrates this load grid for a four-way 2D compound with ROI (top)
and without ROI (bottom).

\subsubsection{ROI in eqPly}
The application should declare its regions of interest during rendering. This
declaration can be very fine-grained, e.g., on each leaf node of a scene
graph. Equalizer will track and optimize multiple regions automatically. The
current implementation of Equalizer merges all regions of a single channel into
one region for compositing and load-balancing. Later versions may use different,
more optimal, heuristics based on the application-declared regions.

Each leaf node of the kd-tree used in eqPly declares its region of interest when
it is rendered. The region is calculated by projecting the bounding box into
screen space and normalizing the resulting screen space rectangle:

{\footnotesize
  \lstinputlisting[firstline=264, lastline=269]
                  {../../../Equalizer/examples/eqPly/vertexBufferLeaf.cpp}
                  ...
  \lstinputlisting[firstline=70, lastline=101]
                  {../../../Equalizer/examples/eqPly/vertexBufferState.cpp}
                  ...
}

The \textsf{declareRegion} is forwarded to \textsf{eq::Channel::declareRegion}.
